This material is based upon work funded by the National Science Foundation Graduate
Fellowship Program, as well as the \gls{US} \gls{DHS} under Grant Award Number
2012-DN-130-NF0001.\footnote{The views and conclusions contained in this
document are those of the author and should not be interpreted as representing
the official policies, either expressed or implied, of the \gls{US} \gls{DHS}.} 

This dissertation would not be possible with the wisdom and patience of my
advisor, Paul Wilson, and the amazing community he fostered, CNERG. Kelly
Burton and Max Lagally both convinced me to remain in graduate school
path---more than once. And Steven ``if you're gonna be dumb you gotta be
tough'' Harrell also kept me in school---more than once, and even after cancer
took him too soon.  

Ni\'{a}:wen'k\'{o}wa to the people who taught me who I am and taught me how to
heal. 

My GERS friends have been with me since the beginning of this journey,
especially Robert, Richard, Chandler, Dinh, Nhi. Extra love to Kalin and
Chelsea, who were there in some dark times and some well-lit ones too.  For
tirelessly encouraging my academic and personal pursuits, I'm appreciative of
my California family, Mel, Bonnie, Joelle, and Jamie.  Thanks to Didier for the
housing while I wrote my dissertation in a pandemic while living nomadically.
Lou, my seestre!  Olivia, my bean sister; thank you for the nitrogen these last
few months.  Heather, thanks for being so supportive of me, for years and
years; I'll come get my mail, I promise.  Robin, you have been such a light in
my life for nearly 15 years, and I know I wouldn't be where I am today without
your friendship and beautiful way of seeing the world. And my last thank you is
to Kreezy, who gracefully tolerate my existence in their apartment even as I
write these acknowledgments, and whose friendship I am so blessed to have.
Konor\'{o}nkwa.

Oht\'{a} akw\'{e}:kon w\'{a}tons!
